
% Thesis Abstract -----------------------------------------------------


%\begin{abstractslong}    %uncommenting this line, gives a different abstract heading
\begin{abstracts}        %this creates the heading for the abstract page

The theory of linear groups goes back to the foundation of group theory. Évariste Galois constructed the groups $PSL(2,p)$ were constructed by Évariste Galois in the 1830s, he also constructed the general linear group over a prime field, $GL(n, p)$, in studying the Galois group of the general equation of degree $p^n$. The existence of non-congruence subgroups was known to Felix Klein who made extensive contribution in the related field of modular forms. This project give's an undergraduate accessible exposition of the theory of congruence and non-congruence subgroups. A goal of the project was to give a more detailed report sized overview of the theory. \\
\\

\textbf{Acknowledgements}: I would like to thank Dr. Kevin Hutchinson for his excellent mentoring and support over the months of the module. I believe his guidance has enabled to do far more than I thought I was capable of and helped me become a much better mathematics student. I would also like to mention Dr. Keith Conrad for his superb collection of expository papers, especially the one on $SL_2(\mathbb{Z})$ \citep{conrad} which served as the initial motivation for beginning this project. \\
I would also like to thank University College Dublin and University of California: Santa Cruz for high quality mathematics education I have been given.
 

\end{abstracts}
%\end{abstractlongs}


% ----------------------------------------------------------------------


%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "../thesis"
%%% End: 
